---
title: "SARS-CoV-2 H-2b epitope predictions"
author: "Hayley McQuilten"
date: "10/10/2020"
output:
  html_document:
    df_print: paged
params:
  data: 'Input/BL-6 MHC I prediction NetMHCpan 4-1.txt'
  strong_binder_EL_threshold: 0.5
  top_binder_EL_threshold: 0.05
  strong_binder_Aff_threshold: 500
  top_binder_Aff_threshold: 50
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
library(ggrepel)
```

## Data import and tidy

```{r Import}
# Import a .txt file copied from NetMHCpan 4.1 output window, tidy data
BL6_MHCI <- read_delim(params$data, 
                       delim = ' ', skip = 18,
                       col_types = cols(
                         Pos = col_double(),
                         MHC = col_character(),
                         Peptide = col_character(),
                         Core = col_character(),
                         Of = col_double(),
                         Gp = col_double(),
                         Gl = col_double(),
                         Ip = col_double(),
                         Il = col_double(),
                         Icore = col_character(),
                         Identity = col_character(),
                         Score_EL = col_double(),
                         `%Rank_EL` = col_double(),
                         Score_BA = col_double(),
                         `%Rank_BA` = col_double(),
                         `Aff(nM)` = col_double(),
                         BindLevel = col_character(),
                         BindStrength = col_character()),
                       comment = '--', trim_ws = T)
```
Warning message expected here, because no data exists in final 2 columns for all entries that are neither strong nor weak binders. So for many rows, expect 18 columns but actual is 16.

Also, the import file contains section breaks throughout the data table. So some rows will contain values that are inconsistent with the specified `col_type` (eg. words or dashes rather than numbers). 

A summary of the parsing issues:
```{r Check 1}
unique(problems(BL6_MHCI)$expected)
```
Should be "18 columns" and "a double".

```{r Check 2}
unique(problems(BL6_MHCI)$actual)
```
Should be "16 columns", then a variety of other words and numbers of columns.

```{r Tidy, echo=FALSE}
tidy <- BL6_MHCI %>%
  filter(!is.na(Pos)) %>%
  separate(Identity, c("Protein", "prefix", "suffix"), sep = "_", extra = "drop") %>%
  select(-BindLevel, -prefix, -suffix) %>%
  mutate(Domain = case_when(Protein == 'S'& Pos >= 334 & Pos <= 518 ~ "RBD",
                            !(Protein == 'S'& Pos >= 334 & Pos <= 518) ~ "Rest of virus",
                            TRUE ~ NA_character_))

# Whole tibble with the proteins in specific order
tidy <- tidy %>%
  mutate(Protein = factor(Protein, levels = c("S", "M", "N", "E", "ORF1ab", "ORF3a",
                                              "ORF6", "ORF7a", "ORF7b", "ORF8", "ORF10")))

# Extra column specifying whether RBD or otherwise
tidy <- tidy %>%
  mutate(Domain = case_when(Protein == 'S'& Pos >= 334 & Pos <= 518 ~ "RBD",
                            !(Protein == 'S'& Pos >= 334 & Pos <= 518) ~ "Rest of virus",
                            TRUE ~ NA_character_))
```


## Plots


```{r Plot, echo=FALSE}
# Subset and plot
# strong binders (top 0.5%) % Rank EL vs BA  
tidy %>%
  filter(`%Rank_EL` <= params$strong_binder_EL_threshold) %>%
  ggplot(aes(`%Rank_EL`, y = `%Rank_BA`)) +
  geom_point(mapping = aes(colour = Protein)) +
  theme_light() +
  scale_color_viridis(discrete=TRUE) +
  labs(title = "Strong binders",
       x = "% Rank eluted ligand (EL)",
       y = "% Rank binding affinity (BA)")
  
# strong binders (top 0.5%, affinity <= 500nM) % Rank EL vs Aff(nM) 
tidy %>%
  filter(`%Rank_EL` <= params$strong_binder_EL_threshold, 
         `Aff(nM)` <= params$strong_binder_Aff_threshold) %>%
  ggplot(aes(`%Rank_EL`, y = `Aff(nM)`)) +
  geom_point(mapping = aes(colour = Protein)) +
  theme_light() +
  scale_color_viridis(discrete=TRUE) +
  labs(title = "Strong binders with affinity <= 500nM",
       x = "% Rank eluted ligand (EL)",
       y = "Affinity (nM)")

# top binders (0.05%)
tidy %>%
  filter(`%Rank_EL` <= params$top_binder_EL_threshold, 
         `Aff(nM)` <= params$top_binder_Aff_threshold) %>%
  ggplot(aes(`%Rank_EL`, y = `Aff(nM)`)) +
  geom_point(mapping = aes(colour = Protein)) +
  theme_light() +
  scale_color_viridis(discrete=TRUE) +
  geom_text_repel(aes(label = Peptide), size = 2.5) +
  labs(title = "Strongest binders",
       x = "% Rank eluted ligand (EL)",
       y = "Affinity (nM)")

# RBD vs rest of virus EL vs BA
tidy %>%
  filter(`%Rank_EL` <= 0.5) %>%
  ggplot(aes(x =`%Rank_EL`, y = `%Rank_BA`)) +
  geom_point(mapping = aes(colour = Domain)) +
  theme_light() +
  scale_color_manual(values=c("#440154FF", "#BEBEBE80")) +
  labs(title = "Highlight on RBD-derived epitopes",
       x = "% Rank eluted ligand (EL)",
       y = "% Rank binding affinity (BA)")

# RBD vs rest of virus EL vs Aff(nM) strongest binders
tidy %>%
  filter(`%Rank_EL` <= params$top_binder_EL_threshold, 
         `Aff(nM)` <= params$top_binder_Aff_threshold) %>%
  ggplot(aes(x =`%Rank_EL`, y = `Aff(nM)`)) +
  geom_point(mapping = aes(colour = Domain)) +
  theme_light() +
  scale_color_manual(values=c("#440154FF", "#BEBEBE80")) +
  labs(title = "Highlight on RBD-derived epitopes, strongest binders",
       x = "% Rank eluted ligand (EL)",
       y = "Affinity (nM)")
```
